---
title: "R Notebook"
output: html_notebook
---

```{r}
library(dplyr)
```


```{r}
DEGs2096 <- read.csv("/scripts/16_CompareKyritsis/Paper/FromEmail/2096 differentially expressed SCI genes.csv")
allDEGs <- read.csv("/scripts/16_CompareKyritsis/Paper/FromEmail/differential_gene_expression_all_contrasts.csv")
outdir <- "/scripts/16_CompareKyritsis/3_DEGscore/1_ProcessLists/outdir/"
```

* HC vs TC = 72 + 160 + 20 + 138 = 390
* HC vs SCI = 1424 + 160 + 20 + 424 = 2028 
* TC vs SCI = 248 + 138 + 20 + 424 = 830

# Step 1: Split up the gene lists so I can recreate the venn

```{r}
# is it the column "TC.HC.adj.P.Val"?
nrow(allDEGs %>% filter(TC.HC.adj.P.Val <= 0.05) %>% filter(logFC > 1 | logFC < -1)) #expect: 390 get: 390
nrow(allDEGs %>% filter(SCI.HC.adj.P.Val <= 0.05) %>% filter(logFC.1 > 1 | logFC.1 < -1)) #expect: 2028 get: 2028
nrow(allDEGs %>% filter(SCI.TC.adj.P.Val <= 0.05) %>% filter(logFC.2 > 1 | logFC.2 < -1)) #expect: 831 get: 830
```

```{r}
TC_HC <- allDEGs %>% filter(TC.HC.adj.P.Val <= 0.05) %>% filter(logFC > 1 | logFC < -1)
SCI_HC <- allDEGs %>% filter(SCI.HC.adj.P.Val <= 0.05) %>% filter(logFC.1 > 1 | logFC.1 < -1)
SCI_TC <- allDEGs %>% filter(SCI.TC.adj.P.Val <= 0.05) %>% filter(logFC.2 > 1 | logFC.2 < -1)
```


# split into upreg vs downreg

```{r}
TC_HC_up <- TC_HC %>% filter(logFC > 0)
TC_HC_down <- TC_HC %>% filter(logFC < 0)

SCI_HC_up <- SCI_HC %>% filter(logFC > 0)
SCI_HC_down <- SCI_HC %>% filter(logFC < 0)

SCI_TC_up <- SCI_TC %>% filter(logFC > 0)
SCI_TC_down <- SCI_TC %>% filter(logFC < 0)
```

```{r}
write.table(TC_HC_up, file = paste0(outdir, "kyritsis_HCvsTC_up.txt"), sep = "\t", quote = FALSE, col.names = NA)
write.table(TC_HC_down, file = paste0(outdir, "kyritsis_HCvsTC_down.txt"), sep = "\t", quote = FALSE, col.names = NA)

write.table(SCI_HC_up, file = paste0(outdir, "kyritsis_HCvsSCI_up.txt"), sep = "\t", quote = FALSE, col.names = NA)
write.table(TC_HC_down, file = paste0(outdir, "kyritsis_HCvsSCI_down.txt"), sep = "\t", quote = FALSE, col.names = NA)

write.table(SCI_TC_up, file = paste0(outdir, "kyritsis_TCvsSCI_up.txt"), sep = "\t", quote = FALSE, col.names = NA)
write.table(SCI_TC_down, file = paste0(outdir, "kyritsis_TCvsSCI_down.txt"), sep = "\t", quote = FALSE, col.names = NA)
```

# now look at the 2096 genes

* A total of 1025 SCI genes upregulated in SCI.TC and SCI.HC but not TC.HC (265 key genes overlap)
* A total of 1071 SCI genes upregulated in SCI.TC and SCI.HC but not TC.HC (159)

```{r}
# all genes (as in the paper)
all2096 <- setdiff((union(SCI_HC$Feature, SCI_TC$Feature)), TC_HC$Feature)

# split by up and downregulated
upreg_2096 <- setdiff((union(SCI_HC_up$Feature, SCI_TC_up$Feature)), TC_HC_up$Feature)
downreg_2096 <- setdiff((union(SCI_HC_down$Feature, SCI_TC_down$Feature)), TC_HC_down$Feature)

# get the subset that is in both SCI_HC and SCI_TC
upreg_2096_key <- setdiff((intersect(SCI_HC_up$Feature, SCI_TC_up$Feature)), TC_HC_up$Feature)
downreg_2096_key <- setdiff((intersect(SCI_HC_down$Feature, SCI_TC_down$Feature)), TC_HC_down$Feature)
```

```{r}
write.table(as.data.frame(all2096), file = paste0(outdir, "kyritsis_all2096.txt"), sep = "\t", quote = FALSE, col.names = NA)

write.table(as.data.frame(upreg_2096), file = paste0(outdir, "kyritsis_2096_upreg.txt"), sep = "\t", quote = FALSE, col.names = NA)
write.table(as.data.frame(downreg_2096), file = paste0(outdir, "kyritsis_2096_downreg.txt"), sep = "\t", quote = FALSE, col.names = NA)

write.table(as.data.frame(upreg_2096_key), file = paste0(outdir, "kyritsis_2096_upreg_KeyOverlap.txt"), sep = "\t", quote = FALSE, col.names = NA)
write.table(as.data.frame(downreg_2096_key), file = paste0(outdir, "kyritsis_2096_downreg_KeyOverlap.txt"), sep = "\t", quote = FALSE, col.names = NA)
```

# what if I make a more streamlined list, i.e. the genes with the top FC

Also filter to require that the genes be included in our dataset

```{r}
inhouse <- readRDS("/scripts/17_MetacellVehicleOnly/vehicle/11.1a_ReplotUMAP/out_obj/ReplotUMAP_Seurat.RDS")
```

```{r}
ProcessGeneList <- function(mylist) {
  mylist_f <- mylist[mylist %in% rownames(inhouse)] #only keep if in inhouse data
  mylist_ft <- allDEGs[allDEGs$Feature %in% mylist_f,] #get the full DE table
  # extract the separate sample data and filter to only keep the significant hits
  mylist_ft_SCI.HC <- mylist_ft[c("Feature", "logFC.1", "SCI.HC.adj.P.Val")] %>%
    filter(SCI.HC.adj.P.Val <= 0.05) %>%
    filter(logFC.1 > 1 | logFC.1 < -1) #extract SCI vs HC data
  colnames(mylist_ft_SCI.HC) <- c("Feature", "logFC", "adjP")
  mylist_ft_SCI.TC <- mylist_ft[c("Feature", "logFC.2", "SCI.TC.adj.P.Val")] %>%
    filter(SCI.TC.adj.P.Val <= 0.05) %>%
    filter(logFC.2 > 1 | logFC.2 < -1) #extract SCI vs TC data
  colnames(mylist_ft_SCI.TC) <- c("Feature", "logFC", "adjP")
  # make separate lists of the up and downregulated genes, and re-merge between samples
  mylist_ft_SCI.HC$comparison <- "SCI.HC"
  mylist_ft_SCI.TC$comparison <- "SCI.TC"
  upreg <- rbind((mylist_ft_SCI.HC %>% filter(logFC > 0)), (mylist_ft_SCI.TC %>% filter(logFC > 0)))
  downreg <- rbind((mylist_ft_SCI.HC %>% filter(logFC < 0)), (mylist_ft_SCI.TC %>% filter(logFC < 0)))
  # remove duplicates
  upreg_f <- upreg %>% arrange(logFC)
  upreg_f <- upreg_f[!duplicated(upreg_f$Feature),]
  downreg_f <- downreg %>% arrange(logFC)
  downreg_f <- downreg_f[!duplicated(downreg_f$Feature),]
  # select top N per direction
  upreg_f$direction <- "up"
  upreg_top50 <- upreg_f %>% arrange(-(logFC)) %>% dplyr::slice(1:50, with_ties = FALSE)
  downreg_f$direction <- "down"
  downreg_top50 <- downreg_f %>% arrange(-(logFC)) %>% dplyr::slice(1:50, with_ties = FALSE)
  all <- rbind(upreg_top50, downreg_top50)
  return(all)
}
```

```{r}
top50_2096 <- ProcessGeneList(all2096)
write.table(top50_2096, file = paste0(outdir, "kyritsis_2096_top50.txt"), sep = "\t", quote = FALSE, col.names = NA)
top50_key <- ProcessGeneList(c(upreg_2096_key, downreg_2096_key))
write.table(top50_key, file = paste0(outdir, "kyritsis_keygenes_top50.txt"), sep = "\t", quote = FALSE, col.names = NA)
```



```{r}
# also pull the 2096 genes
#TC_HC_up[TC_HC_up$Feature %in% DEGs2096$Feature,] #no hits
#TC_HC_down[TC_HC_down$Feature %in% DEGs2096$Feature,] #no hits

#SCI_HC_up_2096 <- SCI_HC_up[SCI_HC_up$Feature %in% DEGs2096$Feature,] #967
#SCI_HC_down_2096 <- SCI_HC_down[SCI_HC_down$Feature %in% DEGs2096$Feature,] #881 hits

#SCI_TC_up_2096 <- SCI_TC_up[SCI_TC_up$Feature %in% DEGs2096$Feature,] #323 hits
#SCI_TC_down_2096 <- SCI_TC_down[SCI_TC_down$Feature %in% DEGs2096$Feature,] #349 hits

# all 2096 up and down
#all2096_up <- as.data.frame(unique(c(SCI_HC_up_2096$Feature, SCI_TC_up_2096$Feature)))
#all2096_down <- as.data.frame(unique(c(SCI_HC_down_2096$Feature, SCI_TC_down_2096$Feature)))
```


