---
title: "R Notebook"
output: html_notebook
---

```{r}
library(Seurat)
library(ggplot2)
library(dplyr)
library(ComplexHeatmap)
library(RColorBrewer)
```


```{r}
kyritsis <- read.csv("/scripts/16_CompareKyritsis/Paper/GSE151371_raw_gene_counts_de-ID.csv", row.names = 1)

inhouse <- readRDS("/scripts/11b_RunClusteringRes0.4/out_obj/allintegratedClustered_Seurat2.RDS")
# filter to keep only blood
keep <- names(which(inhouse$tissue == "blood"))
inhouse <- subset(inhouse, cells = keep)
```

```{r}
outdir <- "/scripts/16_CompareKyritsis/1_LinearRegressionCompare/outdir/"
```

```{r}
# signature genes from table S2
sigGenes <- c("ADAMTS3", "ALDH1A2", "ANKRD34B", "ANO10", "AREG", "ARG1", "ARMC12", "ASPH", "BCL2A1", "C3AR1", "C8orf88", "CAMP", "CARD6", "CCDC183-AS1", "CCM2L", "CDKL5", "CHPT1", "CLEC6A", "CRISP2", "CRISP3", "DACH1", "DHRS9", "DRAM1", "DUSP13", "EFNA1", "FCER1G", "FKBP5", "FMN1", "FOLR3", "FRMD4B", "GABRR2", "GADD45A", "GPR160", "GYG1", "HMGB2", "HN1", "HP", "HPGD", "HTATIP2", "IDI2-AS1", "IFITM3", "IL18RAP", "IL1R2", "ITGA9", "KBTBD7", "KCNE1", "KCNG2", "KL", "LCN2", "LDHA", "LHFP", "LILRB4", "LIN7A", "LINC00671", "LINC01094", "LINC01271", "LINC01581", "LOC100289650", "LOC101927153", "LOC102546294", "LOC642943", "LRRC70", "LTB4R", "MAOA", "MAP2K6", "MB21D1", "MCEMP1", "METTL9", "MILR1", "MIR1273D", "MIR618", "MIR6818", "MMP8", "MS4A4A", "MSRB2", "NANS", "NEDD4", "OLFM4", "OLR1", "OR52W1", "OSBPL1A", "PCOLCE2", "PDSS1", "PLB1", "PLP2", "PPARG", "PPP1R1A", "PRG1", "PRTN3", "PTX3", "RABGEF1", "RNASE1", "RNASE4", "RNF217", "S100P", "SAMSN1", "SAP30", "SCN9A", "SCPEP1", "SDHAF3", "SLC25A24", "SLC38A11", "SLCO4A1", "SNX3", "STBD1", "SYCP2", "TCN1", "TIMP1", "TMCO3", "TMIGD3", "TPST1", "TREML3P", "TTC8", "TTN-AS1", "UBE2J1", "UGCG", "UPB1", "SCI severity  gene ", "A2MP1", "AFAP1", "AHNAK2", "AKAP11", "ALOX12P2", "ARL10", "ASB2", "ATG9B", "ATP10A", "ATP6V0E2", "AXIN2", "BCL9L", "BTBD11", "CACNA1C-AS1", "CACNA1I", "CCDC88C", "CD6", "CD74", "CELSR1", "COLGALT2", "CRLF1", "CTSF", "CTSW", "CXCL10", "DCANP1", "DEPDC7", "DFNB59", "EEF1A1", "ESYT1", "EVL", "FAM102A", "FAM132B", "FAM150B", "FBXL16", "FCRL6", "GFRA2", "GLB1L2", "GZMH", "HLA-DPB1", "HSPG2", "IGSF9B", "KCNA6", "KDF1", "KLHDC7A", "KLRD1", "KNG1", "LANCL1-AS1", "LGR6", "LINC01278", "LOC100507091", "LRRN3", "MLLT6", "MYBPH", "MYCBPAP", "MYO6", "NUP210", "OLFM1", "PASK", "PDGFRB", "PKIA-AS1", "PPP1R16B", "PRSS33", "PTGS2", "RFX7", "RNF125", "RPGRIP1L", "S1PR5", "SARDH", "SCART1", "SERPINA2", "SLC25A23", "SPATA13-AS1", "SPN", "SPOCK2", "TMEM176A", "TMIE", "TTC3P1", "UBASH3A", "UTRN", "ZNF703")
```


Strategy: Log transform each column of data, then calculate the average (for scRNASeq, log transform each cell)

```{r}
# Extract inhouse counts
inhouse_counts <- inhouse@assays$RNA@counts
inhouse_counts <- inhouse_counts[names(which(rowSums(inhouse_counts) > 0)),]

# log transform each column
inhouse_counts_log <- inhouse_counts
inhouse_counts_log <- log2(inhouse_counts_log + 1)
```

```{r}
# remove genes with zero counts
kyritsis <- kyritsis[names(which(rowSums(kyritsis) > 0)),]

# log transform each column in the kyritsis data
kyritsis_log <- kyritsis
kyritsis_log <- log2(kyritsis_log + 1)
```

# Calculate average (log2) gene expression by condition for scRNASeq

```{r}
healthy_cells <- names(which(inhouse$injury == "naive"))
sham_cells <- names(which(inhouse$injury == "sham"))
sci_cells <- names(which(inhouse$injury == "SCI"))

hm_healthy <- inhouse_counts_log[,healthy_cells]
hm_sham <- inhouse_counts_log[,sham_cells]
hm_sci <- inhouse_counts_log[,sci_cells]
```

```{r}
hm_healthy <- as.data.frame(rowMeans(hm_healthy))
hm_sham <- as.data.frame(rowMeans(hm_sham))
hm_sci <- as.data.frame(rowMeans(hm_sci))

colnames(hm_healthy) <- c("hm_healthy")
colnames(hm_sham) <- c("hm_sham")
colnames(hm_sci) <- c("hm_sci")

hm_ave <- cbind(hm_healthy, hm_sham, hm_sci)
```

# For kyritsis, calculate average normalised gene expression by condition

HC = healthy control
TC = sham
SCI = SCI

```{r}
kyr_healthy <- kyritsis_log[,grep("HC", colnames(kyritsis_log))]
kyr_sham <- kyritsis_log[,grep("TC", colnames(kyritsis_log))]
kyr_SCI <- kyritsis_log[,grep("SCI", colnames(kyritsis_log))]

kyr_healthy <- as.data.frame(rowMeans(kyr_healthy))
kyr_sham <- as.data.frame(rowMeans(kyr_sham))
kyr_SCI <- as.data.frame(rowMeans(kyr_SCI))

colnames(kyr_healthy) <- c("kyr_healthy")
colnames(kyr_sham) <- c("kyr_sham")
colnames(kyr_SCI) <- c("kyr_SCI")

kyr_ave <- cbind(kyr_healthy, kyr_sham, kyr_SCI)
```

# merge the tables

```{r}
# genes in both
shared <- intersect(rownames(hm_ave), rownames(kyr_ave))
ave_inhouse_slim <- hm_ave[shared,]
ave_kyr_slim <- kyr_ave[shared,]
# merge
merged_table <- as.data.frame(cbind(ave_inhouse_slim, ave_kyr_slim))
write.table(merged_table, file = paste0(outdir, "MergedAveCounts_Kyritsis+Inhouse.txt"), sep = "\t", quote = FALSE, col.names = NA)
```

# plot for each comparable dataset
Matching celltypes:
* healthy
* sham
* SCI

```{r}
ggplot(merged_table, aes(x = hm_healthy, y = kyr_healthy)) +
  geom_point(color = "black") +
  theme_bw() +
  xlab("naive (in-house)") +
  ylab("naive (Kyritsis)")
ggsave(filename = paste0(outdir, "ExpCompare_", "naive", ".pdf"), width = 5, height = 5)

ggplot(merged_table, aes(x = hm_sham, y = kyr_sham)) +
  geom_point(color = "black") +
  theme_bw() +
  xlab("sham (in-house)") +
  ylab("sham (Kyritsis)")
ggsave(filename = paste0(outdir, "ExpCompare_", "sham", ".pdf"), width = 5, height = 5)

ggplot(merged_table, aes(x = hm_sci, y = kyr_SCI)) +
  geom_point(color = "black") +
  theme_bw() +
  xlab("sham (in-house)") +
  ylab("sham (Kyritsis)")
ggsave(filename = paste0(outdir, "ExpCompare_", "SCI", ".pdf"), width = 5, height = 5)
```

# Same thing, but highlight genes red if they're in the gene signatures list

```{r}
merged_table_inList <- merged_table[rownames(merged_table) %in% sigGenes,]

merged_table_notInList <- merged_table[!(rownames(merged_table) %in% sigGenes),]
```

```{r}
ggplot(merged_table_notInList, aes(x = hm_healthy, y = kyr_healthy)) +
  geom_point(color = "black") +
  geom_point(data = merged_table_inList, aes(x = hm_healthy, y = kyr_healthy), color = "red") +
  theme_bw() +
  xlab("naive (in-house)") +
  ylab("naive (Kyritsis)")
ggsave(filename = paste0(outdir, "ExpCompare_", "naive", "_highlightSignatureGenes.pdf"), width = 5, height = 5)

ggplot(merged_table_notInList, aes(x = hm_sham, y = kyr_sham)) +
  geom_point(color = "black") +
  geom_point(data = merged_table_inList, aes(x = hm_sham, y = kyr_sham), color = "red") +
  theme_bw() +
  xlab("sham (in-house)") +
  ylab("sham (Kyritsis)")
ggsave(filename = paste0(outdir, "ExpCompare_", "sham", "_highlightSignatureGenes.pdf"), width = 5, height = 5)

ggplot(merged_table_notInList, aes(x = hm_sci, y = kyr_SCI)) +
  geom_point(data = merged_table_inList, aes(x = hm_sci, y = kyr_SCI), color = "red") +
  geom_point(color = "black") +
  theme_bw() +
  xlab("SCI (in-house)") +
  ylab("SCI (Kyritsis)")
ggsave(filename = paste0(outdir, "ExpCompare_", "SCI", "_highlightSignatureGenes.pdf"), width = 5, height = 5)
```


```{r}
ggplot(merged_table_inList, aes(x = hm_sci, y = kyr_SCI)) +
  geom_point(color = "red") +
  theme_bw() +
  xlab("SCI (in-house)") +
  ylab("SCI (Kyritsis)")
ggsave(filename = paste0(outdir, "ExpCompare_", "SCI", "_SignatureGenesOnly.pdf"), width = 5, height = 5)
```

```{r}
mycorr_p <- cor(merged_table, method = c("pearson"))
mycorr_s <- cor(merged_table, method = c("spearman"))
```

```{r}
#CellPalette = circlize::colorRamp2(c(0, 0.25, 0.5, 0.75, 1), c("darkblue", "lightskyblue", "white", "lightpink", "darkred"))
CellPalette = circlize::colorRamp2(c(0, 0.25, 0.5, 0.75, 1), c("white", brewer.pal(4, "YlOrRd")))


ht <- Heatmap(mycorr_p,
        cluster_rows = FALSE, cluster_columns = FALSE,
        col = CellPalette, 
        # split up clusters (cols) and additional genes (rows)
        use_raster = TRUE, raster_quality = 1,
        width = unit(5, "cm"),
        height = unit(5, "cm")
        )

pdf(paste0(outdir, "correlation_heatmap.pdf"), height = 6)
draw(ht, annotation_legend_side = "bottom", heatmap_legend_side = "bottom")
dev.off()
```

We are seeing huge batch effect here...

# Calculate linear regression

https://www.datacamp.com/community/tutorials/linear-regression-R

"Not every problem can be solved with the same algorithm. In this case, linear regression assumes that there exists a linear relationship between the response variable and the explanatory variables. This means that you can fit a line between the two (or more variables). In the previous example, it is clear that there is a relationship between the age of children and their height."

```{r}
lm_healthy <- lm(kyr_healthy ~ hm_healthy, data = merged_table)
summary(lm_healthy)
```

Height=a+Age∗b
(AKA: Height = c + Age ∗ m)

In the "Coefficients" section, you can see the values of the intercept (“c” value) and the slope (“m” value) for the age (x). These “c” and “m” values plot a line between all the points of the data. So in this case, if there is a child that is 20.5 months old, c is 64.92 and m is 0.635, the model predicts (on average) that its height in centimeters is around 64.92 + (0.635 * 20.5) = 77.93 cm.

Another aspect to pay attention to your linear models is the p-value of the coefficients. In simple terms, a p-value indicates whether or not you can reject or accept a hypothesis. The hypothesis, in this case, is that the predictor is not meaningful for your model.

The p-value for age is 4.34*e-10 or 0.000000000434. A very small value means that age is probably an excellent addition to your model. The p-value for the number of siblings is 0.85. In other words, there’s 85% chance that this predictor is not meaningful for the regression.

A good way to test the quality of the fit of the model is to look at the RESIDUALS or the differences between the real values and the predicted values. The straight line in the image above represents the predicted values. The red vertical line from the straight line to the observed data value is the residual.

The idea in here is that the sum of the residuals is approximately zero or as low as possible. In real life, most cases will not follow a perfectly straight line, so residuals are expected. In the R summary of the lm function, you can see descriptive statistics about the residuals of the model, following the same example, the red square shows how the residuals are approximately zero.

[in my case, the min is way off 0, and the max is way off, but the 1Q, median and 3Q are approx zero]

Is the data a good fit? One measure very used to test how good is your model is the coefficient of determination or R². This measure is defined by the proportion of the total variability explained by the regression model.

R² = explained variation of the model / total variation of the model

In general, for models that fit the data well, R² is near 1. Models that poorly fit the data have R² near 0. However, it’s essential to keep in mind that sometimes a high R² is not necessarily good every single time (see below residual plots) and a low R² is not necessarily always bad. In real life, events don’t fit in a perfectly straight line all the time. For example, you can have in your data taller or smaller children with the same age. In some fields, an R² of 0.5 is considered good.

In the blue rectangle, notice that there’s two different R², one multiple and one adjusted. The multiple is the R² that you saw previously. One problem with this R² is that it cannot decrease as you add more independent variables to your model, it will continue increasing as you make the model more complex, even if these variables don’t add anything to your predictions (like the example of the number of siblings). For this reason, the adjusted R² is probably better to look at if you are adding more than one variable to the model, since it only increases if it reduces the overall error of the predictions.

Let’s plot the data (in a simple scatterplot) and add the line you built with your linear model. In this example, let R read the data first, again with the read_excel command, to create a dataframe with the data, then create a linear regression with your new data. The command plot takes a data frame and plots the variables on it. In this case, it plots the pressure against the temperature of the material. Then, add the line made by the linear regression with the command abline.

# calculate the stats and replot the plots

```{r}
#lm_DC <- lm(DCs_M ~ DCs, data = merged_table)
lm_healthy <- lm(kyr_healthy ~ hm_healthy, data = merged_table)
#r2 <- round(summary(lm_DC)$r.squared, digits = 2)
ggplot(merged_table, aes(x = hm_healthy, y = kyr_healthy)) +?
  geom_point(color = "black") +
  theme_bw() +
  xlab("healthy (in-house)") +
  ylab("healthy (kyritsis)") +
  geom_smooth(method='lm', se=FALSE, color = "black", size = 0.5) +
  ggpubr::stat_regline_equation(label.x=0, label.y=25) +
  ggpubr::stat_cor(aes(label=..rr.label..), label.x= 0, label.y=20)
ggsave(filename = paste0(outdir, "lm_", "healthy", ".pdf"), width = 5, height = 5)
```

```{r}
#lm_DC <- lm(DCs_M ~ DCs, data = merged_table)
lm_sham <- lm(kyr_sham ~ hm_sham, data = merged_table)
#r2 <- round(summary(lm_DC)$r.squared, digits = 2)
ggplot(merged_table, aes(x = hm_sham, y = kyr_sham)) +
  geom_point(color = "black") +
  theme_bw() +
  xlab("sham (in-house)") +
  ylab("sham (kyritsis)") +
  geom_smooth(method='lm', se=FALSE, color = "black", size = 0.5) +
  ggpubr::stat_regline_equation(label.x=0, label.y=25) +
  ggpubr::stat_cor(aes(label=..rr.label..), label.x= 0, label.y=20)
ggsave(filename = paste0(outdir, "lm_", "sham", ".pdf"), width = 5, height = 5)
```

```{r}
#lm_DC <- lm(DCs_M ~ DCs, data = merged_table)
lm_SCI <- lm(kyr_SCI ~ hm_sci, data = merged_table)
#r2 <- round(summary(lm_DC)$r.squared, digits = 2)
ggplot(merged_table, aes(x = hm_sci, y = kyr_SCI)) +
  geom_point(color = "black") +
  theme_bw() +
  xlab("SCI (in-house)") +
  ylab("SCI (kyritsis)") +
  geom_smooth(method='lm', se=FALSE, color = "black", size = 0.5) +
  ggpubr::stat_regline_equation(label.x=0, label.y=25) +
  ggpubr::stat_cor(aes(label=..rr.label..), label.x= 0, label.y=20)
ggsave(filename = paste0(outdir, "lm_", "SCI", ".pdf"), width = 5, height = 5)
```

# Minh suggests, what if I only plot the top variable genes from the scRNASeq data?

"for most of house keeping genes they appear to be very abundance in high-depth sequencing data"

```{r}
varGenes <- inhouse@assays$integrated@var.features
merged_table_varGenes <- merged_table[rownames(merged_table) %in% varGenes,]
```

merged_table

```{r}
#lm_DC <- lm(DCs_M ~ DCs, data = merged_table)
lm_healthy <- lm(kyr_healthy ~ hm_healthy, data = merged_table_varGenes)
#r2 <- round(summary(lm_DC)$r.squared, digits = 2)
ggplot(merged_table_varGenes, aes(x = hm_healthy, y = kyr_healthy)) +
  geom_point(color = "black") +
  theme_bw() +
  xlab("healthy (in-house)") +
  ylab("healthy (kyritsis)") +
  geom_smooth(method='lm', se=FALSE, color = "black", size = 0.5) +
  ggpubr::stat_regline_equation(label.x=0, label.y=25) +
  ggpubr::stat_cor(aes(label=..rr.label..), label.x= 0, label.y=20)
ggsave(filename = paste0(outdir, "lm_", "healthy", "_variableGenes.pdf"), width = 5, height = 5)
```

```{r}
#lm_DC <- lm(DCs_M ~ DCs, data = merged_table)
lm_sham <- lm(kyr_sham ~ hm_sham, data = merged_table_varGenes)
#r2 <- round(summary(lm_DC)$r.squared, digits = 2)
ggplot(merged_table_varGenes, aes(x = hm_sham, y = kyr_sham)) +
  geom_point(color = "black") +
  theme_bw() +
  xlab("sham (in-house)") +
  ylab("sham (kyritsis)") +
  geom_smooth(method='lm', se=FALSE, color = "black", size = 0.5) +
  ggpubr::stat_regline_equation(label.x=0, label.y=25) +
  ggpubr::stat_cor(aes(label=..rr.label..), label.x= 0, label.y=20)
ggsave(filename = paste0(outdir, "lm_", "sham", "_variableGenes.pdf"), width = 5, height = 5)
```

```{r}
#lm_DC <- lm(DCs_M ~ DCs, data = merged_table)
lm_SCI <- lm(kyr_SCI ~ hm_sci, data = merged_table_varGenes)
#r2 <- round(summary(lm_DC)$r.squared, digits = 2)
ggplot(merged_table_varGenes, aes(x = hm_sci, y = kyr_SCI)) +
  geom_point(color = "black") +
  theme_bw() +
  xlab("SCI (in-house)") +
  ylab("SCI (kyritsis)") +
  geom_smooth(method='lm', se=FALSE, color = "black", size = 0.5) +
  ggpubr::stat_regline_equation(label.x=0, label.y=25) +
  ggpubr::stat_cor(aes(label=..rr.label..), label.x= 0, label.y=20)
ggsave(filename = paste0(outdir, "lm_", "SCI", "_variableGenes.pdf"), width = 5, height = 5)
```