---
title: "R Notebook"
output: html_notebook
---

```{r}
library(Seurat)
library(dplyr)
library(ggplot2)
```

```{r}
inhouse <- readRDS("/scripts/11b_RunClusteringRes0.4/out_obj/allintegratedClustered_Seurat2.RDS")
Idents(inhouse) <- "integrated_snn_res.0.4"
DefaultAssay(inhouse) <- "RNA"
outdir <- "/scripts/16_CompareKyritsis/2_ModuleScore/outdir/"
```

# A simple example of how module score works

```{r}
DimPlot(inhouse, label = TRUE)
```
```{r}
markers <- FindMarkers(inhouse, ident.1 = 6)
markers <- markers %>% filter(avg_log2FC > 0) %>% filter(p_val_adj <= 0.05)
top20 <- markers %>% filter(p_val_adj <= 0.05) %>% arrange(-(avg_log2FC)) %>% dplyr::slice(1:20, with_ties = FALSE)
top20 <- rownames(top20)
list_top20 <- list(top20)
inhouse <- AddModuleScore(inhouse, features = list_top20, assay = "RNA", name = "test_nkt")
FeaturePlot(inhouse, features = "test_nkt1", order = TRUE)
```


# Plot the data

```{r}
DimPlot(inhouse)
```
# Calculate Kyritsis module scores for Table S2

Kyritsis paper is here: `/scripts/16_CompareKyritsis/Paper/jem_20201795.pdf`

Which supp data files do they provide?

Table S2: Genes that change with injury severity. I have reformatted it from a Word document to a .txt file

```{r}
# load the data
goi <- read.delim("/scripts/16_CompareKyritsis/Paper/TableS2_reformat.txt")
up <- goi %>% filter(Direction == "Up") %>% pull(Gene)
down <- goi %>% filter(Direction == "Down") %>% pull(Gene)

# comparison gene set
inhouse_degs <- read.delim("/scripts/13_DEG/AllCellsTogether/4.1_SCIvsshamnaive_blood/outdir_SCIvsNot/markers_SCIvnaive_RemoveRbMt_top100.txt") #genes that are different between SCI and naive in our dataset, across all clusters
```

```{r}
# confirm the genes are present in the data
table(up %in% rownames(inhouse))
table(down %in% rownames(inhouse))
# which genes are missing
up[!(up %in% rownames(inhouse))]
down[!(down %in% rownames(inhouse))]
```


```{r}
up <- list(up)
down <- list(down)
up_inhouse <- list(inhouse_degs %>% filter(direction == "up") %>% pull(gene))
```

```{r}
inhouse <- AddModuleScore(inhouse, features = up, assay = "RNA", name = "score_upreg")
inhouse <- AddModuleScore(inhouse, features = down, assay = "RNA", name = "score_downreg")
inhouse <- AddModuleScore(inhouse, features = up_inhouse, assay = "RNA", name = "score_inhouseSCI")
```

```{r}
FeaturePlot(inhouse, features = "score_upreg1", order = TRUE)
FeaturePlot(inhouse, features = "score_downreg1", order = TRUE)
```

```{r}
FeaturePlot(inhouse, features = "score_upreg1", order = TRUE)
ggsave(filename = paste0(outdir, "feature_up_clust.pdf"), width = 5, height = 5)
FeaturePlot(inhouse, features = "score_downreg1", order = TRUE)
ggsave(filename = paste0(outdir, "feature_down_clust.pdf"), width = 5, height = 5)
FeaturePlot(inhouse, features = "score_inhouseSCI1", order = TRUE)
ggsave(filename = paste0(outdir, "feature_inhouse_clust.pdf"), width = 5, height = 5)

FeaturePlot(inhouse, features = "score_upreg1", order = TRUE, split.by = "sample")
ggsave(filename = paste0(outdir, "feature_up_sample.pdf"), width = 18, height = 3)
FeaturePlot(inhouse, features = "score_downreg1", order = TRUE, split.by = "sample")
ggsave(filename = paste0(outdir, "feature_down_sample.pdf"), width = 18, height = 3)

FeaturePlot(inhouse, features = "score_upreg1", order = TRUE, split.by = "injury")
ggsave(filename = paste0(outdir, "feature_up_injury.pdf"), width = 15, height = 5)
FeaturePlot(inhouse, features = "score_downreg1", order = TRUE, split.by = "injury")
ggsave(filename = paste0(outdir, "feature_down_injury.pdf"), width = 15, height = 5)

FeaturePlot(inhouse, features = "score_upreg1", order = TRUE, split.by = "tissue")
ggsave(filename = paste0(outdir, "feature_up_tissue.pdf"), width = 10, height = 5)
FeaturePlot(inhouse, features = "score_downreg1", order = TRUE, split.by = "tissue")
ggsave(filename = paste0(outdir, "feature_down_tissue.pdf"), width = 10, height = 5)

FeaturePlot(inhouse, features = "score_upreg1", order = TRUE, split.by = "drug")
ggsave(filename = paste0(outdir, "feature_up_drug.pdf"), width = 10, height = 5)
FeaturePlot(inhouse, features = "score_downreg1", order = TRUE, split.by = "drug")
ggsave(filename = paste0(outdir, "feature_down_drug.pdf"), width = 10, height = 5)

FeaturePlot(inhouse, features = "score_upreg1", order = TRUE, split.by = "injurydrug")
ggsave(filename = paste0(outdir, "feature_up_injurydrug.pdf"), width = 16, height = 4)
FeaturePlot(inhouse, features = "score_downreg1", order = TRUE, split.by = "injurydrug")
ggsave(filename = paste0(outdir, "feature_down_injurydrug.pdf"), width = 16, height = 4)
```

```{r}
VlnPlot(inhouse, features = "score_upreg1")
ggsave(filename = paste0(outdir, "vln_up_clust.pdf"), width = 5, height = 5)
VlnPlot(inhouse, features = "score_downreg1")
ggsave(filename = paste0(outdir, "vln_down_clust.pdf"), width = 5, height = 5)
VlnPlot(inhouse, features = "score_inhouseSCI1")
ggsave(filename = paste0(outdir, "vln_inhouse_clust.pdf"), width = 5, height = 5)

VlnPlot(inhouse, features = "score_upreg1", group.by = "sample")
ggsave(filename = paste0(outdir, "vln_up_sample.pdf"), width = 5, height = 5)
VlnPlot(inhouse, features = "score_downreg1", group.by = "sample")
ggsave(filename = paste0(outdir, "vln_down_sample.pdf"), width = 5, height = 5)
VlnPlot(inhouse, features = "score_inhouseSCI1", group.by = "sample")
ggsave(filename = paste0(outdir, "vln_inhouse_sample.pdf"), width = 5, height = 5)

VlnPlot(inhouse, features = "score_upreg1", group.by = "injury")
ggsave(filename = paste0(outdir, "vln_up_injury.pdf"), width = 5, height = 5)
VlnPlot(inhouse, features = "score_downreg1", group.by = "injury")
ggsave(filename = paste0(outdir, "vln_down_injury.pdf"), width = 5, height = 5)
VlnPlot(inhouse, features = "score_inhouseSCI1", group.by = "injury")
ggsave(filename = paste0(outdir, "vln_inhouse_injury.pdf"), width = 5, height = 5)

VlnPlot(inhouse, features = "score_upreg1", group.by = "tissue")
ggsave(filename = paste0(outdir, "vln_up_tissue.pdf"), width = 5, height = 5)
VlnPlot(inhouse, features = "score_downreg1", group.by = "tissue")
ggsave(filename = paste0(outdir, "vln_down_tissue.pdf"), width = 5, height = 5)
VlnPlot(inhouse, features = "score_inhouseSCI1", group.by = "tissue")
ggsave(filename = paste0(outdir, "vln_inhouse_tissue.pdf"), width = 5, height = 5)

VlnPlot(inhouse, features = "score_upreg1", group.by = "drug")
ggsave(filename = paste0(outdir, "vln_up_drug.pdf"), width = 5, height = 5)
VlnPlot(inhouse, features = "score_downreg1", group.by = "drug")
ggsave(filename = paste0(outdir, "vln_down_drug.pdf"), width = 5, height = 5)
VlnPlot(inhouse, features = "score_inhouseSCI1", group.by = "drug")
ggsave(filename = paste0(outdir, "vln_inhouse_drug.pdf"), width = 5, height = 5)

VlnPlot(inhouse, features = "score_upreg1", group.by = "injurydrug")
ggsave(filename = paste0(outdir, "vln_up_injurydrug.pdf"), width = 5, height = 5)
VlnPlot(inhouse, features = "score_downreg1", group.by = "injurydrug")
ggsave(filename = paste0(outdir, "vln_down_injurydrug.pdf"), width = 5, height = 5)
VlnPlot(inhouse, features = "score_inhouseSCI1", group.by = "injurydrug")
ggsave(filename = paste0(outdir, "vln_inhouse_injurydrug.pdf"), width = 5, height = 5)
```

```{r}
VlnPlot(inhouse, features = "score_upreg1", split.by = "sample")
ggsave(filename = paste0(outdir, "vlnBySample_up_sample.pdf"), width = 8, height = 5)
VlnPlot(inhouse, features = "score_downreg1", split.by = "sample")
ggsave(filename = paste0(outdir, "vlnBySample_down_sample.pdf"), width = 8, height = 5)
VlnPlot(inhouse, features = "score_inhouseSCI1", split.by = "sample")
ggsave(filename = paste0(outdir, "vlnBySample_inhouse_sample.pdf"), width = 8, height = 5)

VlnPlot(inhouse, features = "score_upreg1", split.by = "injury")
ggsave(filename = paste0(outdir, "vlnBySample_up_injury.pdf"), width = 8, height = 5)
VlnPlot(inhouse, features = "score_downreg1", split.by = "injury")
ggsave(filename = paste0(outdir, "vlnBySample_down_injury.pdf"), width = 8, height = 5)
VlnPlot(inhouse, features = "score_inhouseSCI1", split.by = "injury")
ggsave(filename = paste0(outdir, "vlnBySample_inhouse_injury.pdf"), width = 8, height = 5)

VlnPlot(inhouse, features = "score_upreg1", split.by = "tissue")
ggsave(filename = paste0(outdir, "vlnBySample_up_tissue.pdf"), width = 8, height = 5)
VlnPlot(inhouse, features = "score_downreg1", split.by = "tissue")
ggsave(filename = paste0(outdir, "vlnBySample_down_tissue.pdf"), width = 8, height = 5)
VlnPlot(inhouse, features = "score_inhouseSCI1", split.by = "tissue")
ggsave(filename = paste0(outdir, "vlnBySample_inhouse_tissue.pdf"), width = 8, height = 5)

VlnPlot(inhouse, features = "score_upreg1", split.by = "drug")
ggsave(filename = paste0(outdir, "vlnBySample_up_drug.pdf"), width = 8, height = 5)
VlnPlot(inhouse, features = "score_downreg1", split.by = "drug")
ggsave(filename = paste0(outdir, "vlnBySample_down_drug.pdf"), width = 8, height = 5)
VlnPlot(inhouse, features = "score_inhouseSCI1", split.by = "drug")
ggsave(filename = paste0(outdir, "vlnBySample_inhouse_drug.pdf"), width = 8, height = 5)

VlnPlot(inhouse, features = "score_upreg1", split.by = "injurydrug")
ggsave(filename = paste0(outdir, "vlnBySample_up_injurydrug.pdf"), width = 8, height = 5)
VlnPlot(inhouse, features = "score_downreg1", split.by = "injurydrug")
ggsave(filename = paste0(outdir, "vlnBySample_down_injurydrug.pdf"), width = 8, height = 5)
VlnPlot(inhouse, features = "score_inhouseSCI1", split.by = "injurydrug")
ggsave(filename = paste0(outdir, "vlnBySample_inhouse_injurydrug.pdf"), width = 8, height = 5)
```

# plot the scores for my "up_inhouse" list as a comparison

```{r}

```

# Module 13

```{r}
mod13 <- list(c("TMED8",
"GYG1",
"ZNF438",
"TMCO3",
"YIPF1",
"KLHL2",
"ANXA3",
"CYSTM1",
"CLEC4D",
"BMX",
"ST6GALNAC3",
"NLRC4",
"S100A12",
"RABGEF1",
"SLC37A3"))

inhouse <- AddModuleScore(inhouse, features = mod13, assay = "RNA", name = "module13")
FeaturePlot(inhouse, features = "module131", order = TRUE)
```
```{r}
VlnPlot(inhouse, features = "module131")
ggsave(filename = paste0(outdir, "vln_", sampleID, "_clust.pdf"), width = 5, height = 5)

VlnPlot(inhouse, features = "module131", group.by = "sample")
ggsave(filename = paste0(outdir, "vln_", sampleID, "_sample.pdf"), width = 5, height = 5)

VlnPlot(inhouse, features = "module131", group.by = "injury")
ggsave(filename = paste0(outdir, "vln_", sampleID, "_injury.pdf"), width = 5, height = 5)

VlnPlot(inhouse, features = "module131", group.by = "tissue")
ggsave(filename = paste0(outdir, "vln_", sampleID, "_tissue.pdf"), width = 5, height = 5)

VlnPlot(inhouse, features = "module131", group.by = "drug")
ggsave(filename = paste0(outdir, "vln_", sampleID, "_drug.pdf"), width = 5, height = 5)

VlnPlot(inhouse, features = "module131", group.by = "injurydrug")
ggsave(filename = paste0(outdir, "vln_", sampleID, "_injurydrug.pdf"), width = 5, height = 5)

VlnPlot(inhouse, features = "module131", split.by = "sample")
ggsave(filename = paste0(outdir, "vlnBySample_", sampleID, "_sample.pdf"), width = 8, height = 5)

VlnPlot(inhouse, features = "module131", split.by = "injury")
ggsave(filename = paste0(outdir, "vlnBySample_", sampleID, "_injury.pdf"), width = 8, height = 5)

VlnPlot(inhouse, features = "module131", split.by = "tissue")
ggsave(filename = paste0(outdir, "vlnBySample_", sampleID, "_tissue.pdf"), width = 8, height = 5)

VlnPlot(inhouse, features = "module131", split.by = "drug")
ggsave(filename = paste0(outdir, "vlnBySample_", sampleID, "_drug.pdf"), width = 8, height = 5)

VlnPlot(inhouse, features = "module131", split.by = "injurydrug")
ggsave(filename = paste0(outdir, "vlnBySample_", sampleID, "_injurydrug.pdf"), width = 8, height = 5)
```

```{r}
mod12 <- list(c("MMP8",
"LTF",
"LCN2",
"BPI",
"ITGA9",
"CRISP3",
"CEACAM6",
"CHIT1",
"CEACAM8",
"TCN1",
"ERG",
"MPO",
"OLR1",
"OLFM4",
"RGL4"
))

mod1 <- list(c("FAM168B",
"OTULIN",
"TAF1",
"HNRNPD",
"HNRNPUL2",
"QRICH1",
"TRAPPC10",
"TNRC6A",
"SMARCA2",
"RALGAPB",
"TSPYL1",
"SART3",
"YLPM1",
"NSD1",
"USP7"
))

mod5 <- list(c("GOLGA3",
"SMYD4",
"NOL9",
"EXOSC6",
"MFATC2IP",
"TNRC6CAS1",
"PEX26",
"POLH",
"CYP20A1",
"MRI1",
"GNL3L",
"FOXK1",
"POU6F1",
"NF2",
"CNNM3"
))

mod10 <- list(c("HERC5",
"IFIT3",
"IFIH1",
"PARK14",
"IFIT2",
"IFIT1",
"STAT1",
"XAF1",
"NUB1",
"GBP1",
"IFI44L",
"ZCCHC2",
"DDX58",
"HERC6",
"SP140"
))

mod16 <- list(c("C14orf2",
"TIMM8b",
"COX5B",
"COX7A2",
"NDUFA1",
"COX7B",
"TMEM258",
"SDHB",
"MYL6",
"COX6A1",
"TXNDC17",
"NDUFB1",
"RNF181",
"UQCR10",
"PSMB3"
))
```

```{r}
inhouse <- AddModuleScore(inhouse, features = mod12, assay = "RNA", name = "module12_")
inhouse <- AddModuleScore(inhouse, features = mod1, assay = "RNA", name = "module1_")
inhouse <- AddModuleScore(inhouse, features = mod5, assay = "RNA", name = "module5_")
inhouse <- AddModuleScore(inhouse, features = mod10, assay = "RNA", name = "module10_")
inhouse <- AddModuleScore(inhouse, features = mod16, assay = "RNA", name = "module16_")

FeaturePlot(inhouse, features = "module12_1", order = TRUE)
FeaturePlot(inhouse, features = "module1_1", order = TRUE)
FeaturePlot(inhouse, features = "module5_1", order = TRUE)
FeaturePlot(inhouse, features = "module10_1", order = TRUE)
FeaturePlot(inhouse, features = "module16_1", order = TRUE)
```

```{r}
plotScores <- function(featureID, sampleID) {
VlnPlot(inhouse, features = featureID)
ggsave(filename = paste0(outdir, "vln_", sampleID, "_clust.pdf"), width = 5, height = 5)

VlnPlot(inhouse, features = featureID, group.by = "sample")
ggsave(filename = paste0(outdir, "vln_", sampleID, "_sample.pdf"), width = 5, height = 5)

VlnPlot(inhouse, features = featureID, group.by = "injury")
ggsave(filename = paste0(outdir, "vln_", sampleID, "_injury.pdf"), width = 5, height = 5)

VlnPlot(inhouse, features = featureID, group.by = "tissue")
ggsave(filename = paste0(outdir, "vln_", sampleID, "_tissue.pdf"), width = 5, height = 5)

VlnPlot(inhouse, features = featureID, group.by = "drug")
ggsave(filename = paste0(outdir, "vln_", sampleID, "_drug.pdf"), width = 5, height = 5)

VlnPlot(inhouse, features = featureID, group.by = "injurydrug")
ggsave(filename = paste0(outdir, "vln_", sampleID, "_injurydrug.pdf"), width = 5, height = 5)

VlnPlot(inhouse, features = featureID, split.by = "sample")
ggsave(filename = paste0(outdir, "vlnBySample_", sampleID, "_sample.pdf"), width = 8, height = 5)

VlnPlot(inhouse, features = featureID, split.by = "injury")
ggsave(filename = paste0(outdir, "vlnBySample_", sampleID, "_injury.pdf"), width = 8, height = 5)

VlnPlot(inhouse, features = featureID, split.by = "tissue")
ggsave(filename = paste0(outdir, "vlnBySample_", sampleID, "_tissue.pdf"), width = 8, height = 5)

VlnPlot(inhouse, features = featureID, split.by = "drug")
ggsave(filename = paste0(outdir, "vlnBySample_", sampleID, "_drug.pdf"), width = 8, height = 5)

VlnPlot(inhouse, features = featureID, split.by = "injurydrug")
ggsave(filename = paste0(outdir, "vlnBySample_", sampleID, "_injurydrug.pdf"), width = 8, height = 5)
}
```

```{r}
plotScores("module1_1", "mod1")
plotScores("module10_1", "mod10")
plotScores("module12_1", "mod12")
plotScores("module16_1", "mod16")
plotScores("module5_1", "mod5")
```

