---
title: "11.5_NetworkToUMAP.Rmd"
author: "Laura Grice"
date: "9th July 2020"
output: html_notebook
---

TO DO: UPDATE TO SELECT METACELL NODES, AS PER THIS FILE: /scripts/2_output/11_metacell/11.6_NetworkToUmap/DC/select_by_metacell/DC_11.6_NetworkToUMAP_byMetacell.Rmd














Goal: To select a region of interest in a Metacell network diagram, extract the cells associated with those metacells/that region, and visualise them in a Seurat UMAP plot.

# User input

Here:
* SampleID = some name (can be anything - but it's good if it's something informative about the region you will pick, e.g. "neut_bottom"
* seurat.path = path to the Seurat object you want to map cells to
* seurat.annotation.col = name of a Seurat metadata column (will be used to colorise the original network diagram)
* outdir_metacell_figs and _objs = the directories where the Metacell analysis was originally run
* workdir = any directory, where output from this re-analysis will be saved

```{r}
sampleID <- "neut_bottom"
seurat.path <- "/scripts/2_output/10i_ClusterContaminants/out_obj/df_subclusters.RDS"
seurat.annotation.col <- "subclust"
outdir_metacell_figs <- "/scripts/2_output/11_metacell/11.3_RunMetacell/neut/out_fig/"
outdir_metacell_obj <- "/scripts/2_output/11_metacell/11.3_RunMetacell/neut/out_obj/"
workdir <- "/scripts/1a_sandbox/20200708_sandbox_markersOnNetwork/out/"
```

```{r}
library(metacell)
library(Seurat)
library(ggplot2)
library(gatepoints)
```

# Visualise the existing data

## Plot the network diagram

```{r}
# Re-initialise the Metacell data
scdb_init(outdir_metacell_obj, force_reinit = T)
scfigs_init(outdir_metacell_figs)

# Not strictly necessary - change the analysis ID so that any changes to the Metacell objects don't over-write the original analysis
mc_play = scdb_mc("test_mc_f")
scdb_add_mc("playing", mc_play)

# Prepare annotations to colourise the network diagram
## first get the order of cells we need to colour
mc2d = scdb_mc2d("test_2dproj")
cellnames <- names(mc2d@sc_x)
mc = scdb_mc(mc2d@mc_id)
## get the cellID-to-cluster conversion, and filter it so it is the same length and order as the cellID-to-metacell conversion
df <- readRDS(seurat.path)
clusterannotations <- df[[seurat.annotation.col]]
clusterannotations <-clusterannotations[cellnames, , drop = FALSE]
## get the node and edge data
fr = mc2d@graph$mc1
to = mc2d@graph$mc2
```

```{r}
# Build the network diagram
## first need to merge all the data into a table
networkdata <- cbind(as.data.frame(mc2d@sc_x), as.data.frame(mc2d@sc_y), clusterannotations)
colnames(networkdata) <- c("mc2d_xval", "mc2d_yval", "cluster")
## make a custom palette
length <- length(unique(networkdata$cluster)) + 2 #longer, i don't like the first two colours
palette <- DiscretePalette(length, palette = "polychrome")
palette <- palette[3:length]
## plot the graph
network <- ggplot() + 
  geom_point(aes(x=networkdata$mc2d_xval, y=networkdata$mc2d_yval, color=networkdata$cluster), size = 1) +
  scale_color_manual(values = palette) +
  geom_segment(aes(x = mc2d@mc_x[fr], y = mc2d@mc_y[fr], xend = mc2d@mc_x[to], yend = mc2d@mc_y[to]), col = "gray43") +
  geom_point(aes(x = mc2d@mc_x, y = mc2d@mc_y), shape = 21, fill = "lightyellow", color = "black", size = 6) +
  geom_text(aes(x = mc2d@mc_x, y = mc2d@mc_y, label = 1:length(mc2d@mc_x)), size = 3.5) +
  theme_void() +
  theme(legend.position = "bottom")
network
#ggsave(filename = paste0(workdir, sampleID, "_networkdiagram.pdf"), width = 10, height = 10)
```

## Plot the Seurat dim plot
```{r}
## get axis range
plot <- DimPlot(df)
xrange <- ggplot_build(plot)$layout$panel_scales_x[[1]]$range$range
yrange <- ggplot_build(plot)$layout$panel_scales_y[[1]]$range$range

df2 <- subset(df, cells = cellnames)
DimPlot(df2, group.by = seurat.annotation.col, label = TRUE, repel = TRUE, cols = palette) +
   NoLegend() +
   xlim(xrange) +
   ylim(yrange)
# ggsave(paste0(workdir, sampleID, "_seuratplot.pdf"))
```

# Select the points of interest

## Prepare the data first

```{r fig.height=8, fig.width=8}
data <- networkdata
remove <- which(is.na(data$mc2d_xval) | is.na(data$mc2d_yval))
if (length(remove) > 0) {
data <- data[-remove,]  
}
data$cluster <- droplevels(data$cluster)
plot(x=data$mc2d_xval, y=data$mc2d_yval, col = palette[ unclass(data$cluster) ], cex = 0.5, pch = 16)
```

try and plot better colours
  geom_point(aes(x=networkdata$mc2d_xval, y=networkdata$mc2d_yval, color=networkdata$cluster), size = 1) +


```{r}
# call the following in the console:
#plot(x=data$mc2d_xval, y=data$mc2d_yval, col = palette[ unclass(data$cluster) ], cex = 0.5, pch = 16)
#selectedPoints <- fhs(data, mark = TRUE)
```

## Re-plot the graph, highlighting your selected points

Sanity check.

```{r}
selectedPoints <- as.vector(selectedPoints)
# colour as red
colours <- ifelse(rownames(data) %in% selectedPoints, "red", "black")
plot(data[,1:2], col = colours, cex = 0.5, pch = 16)
```

# Use as metadata in your Seurat object

Note, here we use a Seurat object containing only the cells used in Metacell, but it won't matter if you want to use the full Seurat object.

```{r}
meta <- as.data.frame(rownames(df2[[]]))
colnames(meta) <- "cells"
meta <- mutate(meta, "selected" = ifelse(meta$cells %in% selectedPoints, "selected", "not"))
rownames(meta) <- meta$cells
meta$cells <- NULL
```

```{r}
df2 <- AddMetaData(df2, metadata = meta)
```

```{r fig.height=5, fig.width=5}
DimPlot(df2, group.by = "selected", order = c("selected"))
```

```{r}
table <- as.data.frame(selectedPoints)
write.table(table, paste0(workdir, sampleID, "_ROI.txt"), quote = FALSE, sep = "\t", col.names = FALSE, row.names = FALSE)
```

